---
title: Installation
toc: false
---



The easiest way to install `repsim` is through official repositories such as **PyPI** and **CRAN**.


::: {.panel-tabset group="language"}
### Python
```{bash}
#| echo: true
#| eval: false
pip install -U repsim
```

### R
```{r}
#| echo: true
#| eval: false
install.packages("repsim")
```
:::

The development version can also be installed from source hosted on GitHub:


::: {.panel-tabset group="language"}
### Python
```{bash}
#| echo: true
#| eval: false
pip install git+https://github.com/kisungyou/repsim.git#subdirectory=repsim-Python
```

### R
```{r}
#| echo: true
#| eval: false
remotes::install_github("kisungyou/repsim", subdir = "repsim-R")
```
:::

### Dependencies

`repsim` is built on a shared `C++17` computational backend, ensuring consistent performance across Python and R. To install and use `repsim`, your system should have 

- A `C++17`-compatible compiler (e.g., GCC ≥ 9, Clang ≥ 10, or MSVC ≥ 2019)
- CMake ≥ 3.20 (required only when building from source)
- Standard build tools available on Linux, macOS, and Windows

For language-specific requirements, see below:

::: {.panel-tabset group="language"}
### Python

Requires a modern Python environment (≥ 3.8) and the following build dependencies:

- `scikit-build-core` ≥ 0.10 - build backend integrating CMake with PEP 517
- `pybind11` ≥ 2.11 -Python/C++ interface
- `numpy` ≥ 1.23 - array and numerical operations
- `packaging` ≥ 23.0 - version and metadata management

### R

Requires R ≥ 4.2 and a working `C++17` toolchain. All necessary C++ headers (`Rcpp`, `Eigen`) are bundled and linked automatically and no additional R dependencies need to be installed manually.
:::


