---
title: "This is fake"
description: "Centered Kernel Alignment similarity between multiple representations."
sidebar: api
toc: true
categories: ["Kernelized"] 
---

```{r}
#| include: false
library(reticulate)
# Point reticulate to the exact Python in your conda env
use_python("/Users/kyou/miniconda3/envs/repsim/bin/python", required = TRUE)
py_config()  # sanity print (will show the Python path)
```



Dot-product similarity between multiple representations (Kornblith et al., 2019).

::: {.panel-tabset}
## Python

**Function:** `repsim.dot_product(mats)`

### Parameters
- **mats**: list or tuple of M arrays, each with shape `(n_samples, n_features_k)`.  
  All arrays must have the same number of rows (samples).

### Returns
- **ndarray** of shape `(M, M)` — symmetric matrix of pairwise dot-product similarities.

### Example

```{python}
import numpy as np, repsim as rs
X1, X2 = np.random.randn(100, 32), np.random.randn(100, 64)
rs.dot_product([X1, X2])
```

---

## R

**Function:** `dot_product(mats)`


### Parameters
- **mats**: list of M numeric matrices, each with dimension `(n_samples, n_features_k)` and identical row counts.

### Returns
- **matrix** of size `(M,M)` containing pairwise dot-product similarities.


### Example
```{r}
library(repsim)
set.seed(42)
X1 <- matrix(rnorm(100*32), 100, 32)
X2 <- matrix(rnorm(100*64), 100, 64)
dot_product(list(X1, X2))
```
:::